---
title: "Data Visualization for Data Exploration"
author: "Edwin Maldonado - `emaldonado1127@floridapoly.com`"
output: html_notebook
---

### Loading libraries


```{r}
library("dplyr") 
library("ggplot2")
library("tidyverse")
```

### Reading the dataset

```{r}
vehicle <- read_csv("C:\\Users\\maldo\\OneDrive\\Desktop\\FloridaPoly\\Data vizualization\\dataviz_mini-project_01 Edwin Maldonado\\data\\fuel.csv", col_types = cols())
```

### Checking dataset

```{r}
head(vehicle)
```

```{r}
str(vehicle)
```

Let's do some grouping.

First, let's check how many regular fuel car types are available in this data set.

```{r}
vehicle1 <- vehicle %>%
  group_by(fuel_type) %>%
  summarize(total = n()) %>% 
  mutate(proportion = total / sum(total))

vehicle1
```

## What is the most common type of automobile fuel?

Let's visualize this chart into a lollipop graph.

```{r}

P1.1 <- ggplot(vehicle1,aes(x = fuel_type, y = proportion, color = fuel_type)) +
  
        geom_pointrange(aes(ymin = 0, ymax = proportion)) +
        guides(color = FALSE) +
  
        labs(x = "", y = "Proportion", title = "Quantity of various types of car fuels", subtitle = "Vehicles") +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=1, family="Courier", face="italic"),
              plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
              plot.subtitle = element_text(size = 12, hjust = 0.5))
P1.1
```

```{r}
install.packages("svglite")
```

```{r}
library("svglite")
ggsave("Quantity of various types of car fuels.jpg", P1.1, width = 4, height = 4)

```


_**Comment**_

As we see in this graph, the most common fuel type is "regular". More than 60% of cars use this type of fuel. We normally know this fuel as "gasoline".

I was specting this result since this is a common fuel, but I was also wondering about how far we are from renewable fuel.

For this first visualization, we just consider the "Fuel_type" column. There are other columns that need to be included, like "fuel_type_1".


## How efficient is performance in cities and highways by fuel type?.

```{r}

P1.2 <- ggplot(vehicle, aes(composite_city_mpg, composite_highway_mpg, color = fuel_type)) +
   
        geom_jitter(width = .5, size=3) +
   
       labs(subtitle="(mpg)", 
                   y="Highway", 
                   x="City", 
               title="Efficient by fuel type",
              color = "Fuel type") +
       theme(legend.position="right",
              plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
              plot.subtitle = element_text(size = 12, hjust = 0.5))

P1.2 
```

```{r}
ggsave("Efficient by fuel type.jpg", P1.2, width = 7, height = 4)
```

_**Comment**_

We can see that electric automobiles and fuels derived from non-renewable resources have the maximum efficiency.
This is a frequent graphic that we make with this type of data set, but I chose to look at this type of fuel since we are paying more for it these days.

Along with efficiency, cost will be an important factor to consider. That may be a fascinating investigation.


## How fuels type have been implemented over time?

```{r}
library("ggridges")
P1.3 <- ggplot(vehicle, aes(x = year, y = fuel_type, fill = fuel_type)) +
  geom_density_ridges() +
  guides(fill = FALSE)+
  scale_fill_viridis_d(option = "m")+
  labs(y = "", x = "Years", title = "Implemented fuels over time", subtitle = "Vehicles")+
  theme(
        legend.position = "none",
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, hjust = 0.5))

P1.3
```

```{r}
ggsave("Implemented fuels over time.jpg", P1.3, width = 7, height = 4)
```

_**Comment**_

It can be interesting to plot some historical data about this fuel introduction. In this plot, there is a decrease in regular fuel. Electric fuel or cars are getting more attention. This may be because of preview analysis about efficacy.

It is important to point out that companies are looking to replace nonrenewable fuel for renewable fuel. This plot confirms this assumption.
